### 2019-05-06 paper-four/paper-four.tex
paper four first post

\documentclass{article}

\usepackage{enumitem}

\setlength{\parindent}{1.3em}
\setlength{\parskip}{0.7em}
\setlist{noitemsep}

\begin{document}

\begin{titlepage}
   \vspace*{\stretch{1.0}}
   \begin{center}
     \Huge\textbf{Vipassana for Hackers}\\
     \Huge{Paper Four: Mastering Spacetime}\\
     \vspace{5cm}
     \large\textit{Steven Deobald}\\
     \large\textit{Version 0.1 DRAFT}\\
     \large\textit\today\\
     \vspace{5cm}
     \large\textit{}\\
   \end{center}
   \vspace*{\stretch{2.0}}
\end{titlepage}

\begin{center}
  \Huge{Target Audience}
\end{center}

\end{document}